---
title: "R homework"
author: "Tóth-Király István"
date: "June 12, 2018"
output: html_document
---
# Abstract
Given the current Migrant crisis throughout Europe, the following study examined whether gender tendency of dehumanization is related to prejudice-related outcomes of modern racism and xenophobia. Additionally, it was also examined whether group malleability (i.e., the belief that members of a group can change) can mediate these relationships. For this purpose, a nationally representative sample of 505 Hungarians were recruited and completed measures related to dehumanization, group malleability, racism, and xenophobia. The results provided support for the proposed model: dehumanization directly predicted malleability, racism, and xenophobia, while it also indirectly predicted the two outcomes through group malleability. These results suggest that future intervention studies could target the reduction of prejudice by addressing one's beliefs about the changeability of groups.

# Hypotheses
1. It was assumed that general tendency of dehumanization is negatively related to group malleability beliefs. If one perceives social groups with less human qualities, this person will think that the given group cannot be changed.
2. It was also expected that group malleability beliefs mediates between dehumanization and prejudice outcome measures of racism towards Migrants and general xenophobia. More specifically, it was expected that the higher level of dehumanization will lead to stronger beliefs in unchangeability of the groups that finally leads to higher level of racism, xenophobia, and threat of outgroups.

# Sample
This research employed a nationally representative probability sample of Hungarians who used the Internet at least once a week. Participants were selected randomly from 15,000 panel members with the help of a research market company in June of 2016. For the preparation of the sample, a multiple-step, proportionally stratified, probabilistic sampling method was employed, in which individuals were removed from the panel if they gave responses too quickly (i.e., without paying attention to their response) and/or had fake (unused) e-mail addresses (see Tóth-Király, Bothe, & Orosz, 2017 for similar procedures). The final sample of 505 Hungarian respondents who gave valid answers was nationally representative-among those who use Internet at least once a week-in terms of gender (female = 258; 51.1%), age (Mage = 40.19 years; SDage = 11.79 years; range 18-60 years), education (18% had primary level of education, 33% graduated from high school and 48.9% had higher education degree) and place of residence (18.6% in capital city, 21.4% in county capitals, 31.5% in towns and 28.5% in villages).

# Necessary packages 
```{r}
library(haven)      # for data importation
library(tidyverse)  # for data manipulation
library(MVN)        # for testing normality
library(MBESS)      # for calculating model-based omega composite reliability
library(psych)      # for calculating Cronbach's alpha and correlations
library(lavaan)     # for structural equation modeling
```

# Preliminary analyses

### Data importation
```{r}
dehumi <- read_spss("dehumi.sav")
```

### Variable calculation
```{r}
variables <-
  dehumi %>%
  mutate(east_avg = (syr + turk + lib)/3,
         west_avg = (ger + dane + usa)/3,
         hung_avg = (hun + hunw + hune)/3,
         dehu_avg = (east_avg + west_avg + hung_avg)/3,
         mall_avg = (mall1 + mall2 + mall3 + mall4)/4,
         raci_avg = (rac1 + rac2 + rac3 + rac4 + rac5 + rac6 + rac7)/7,
         xeno_avg = (xeno1 + xeno2 + xeno3 + xeno4 + xeno5)/5)
```
* General tendency of dehumanization (Orosz et al., in prep.): higher scores reflect lower dehumanization
* Group malleability (HAlperin et al., 2011): higher scores reflect more fixed and stable attitudes towards groups (i.e., they cannot change)
* Modern racism (McConahay, 1986): higher scores reflect more racist attitudes
* Xenophobia (van der Veer et al., 2011): higher scores reflect more xenophobic attitudes

### Assessment of univariate and multivariate normality with skewness/kurtosis & Mardia's test
```{r}
onlygood <- variables %>%
  select(dehu_avg:xeno_avg)

mvn(onlygood)
```

As the variables did not have normality, Spearman's rank-order correlation was used for assessing the relationships between the variables and the robust maximum-likelihood (MLR) estimator was used in SEM to account for the non-normality.

### Correlation between the variables
```{r}
  corr.test(onlygood, method = "spearman")  %>%
  print(short = FALSE)
```
The examined correlations between the variables are reasonable in direction and magnitude.

# Measurement models and reliabities
The constructs are evaluated on the basis of conventional fit indices: comparative fit index (CFI), Tucker-Lewis index (TLI), and the root mean square error of approximation (RMSEA) with its 90% confidence interval (CI). According to typical interpretation guidelines (Hu & Bentler, 1999; Marsh et al., 2005), values greater than .90 or .95 for CFI and TLI indicate satisfactory or excellent fit, respectively. Alternatively, for RMSEA, values below .08 are considered acceptable, while values below .06 are considered excellent.

### General tendency of dehumanization
```{r}
dehuman <- '
  tend =~ east + west + hung
  east =~ syr + turk + lib
  west =~ ger + dane + usa
  hung =~ hun + hunw + hune '
fit_dehum <- cfa(dehuman, data = dehumi, estimator = 'MLR')
summary(fit_dehum, fit.measures = TRUE, standardized = TRUE)
```
The fit of this scale is adequate.

Given that Cronbach's alpha has been found to be less reliable (Rodriguez et al., 2016), omega coefficients were also calculated.

Alpha for dehumanization
```{r}
variables %>%
  select(syr:hune) %>%
  alpha()
```
Omega total for dehumanization
```{r}
variables %>%
  select(syr:hune) %>%
  ci.reliability()
```

### Group malleability
```{r}
malleab <- 'mall =~ mall1 + mall2 + mall3 + mall4 '
fit_malleab <- cfa(malleab, data = dehumi, estimator = 'MLR')
summary(fit_malleab, fit.measures = TRUE, standardized = TRUE)
```
While CFI and TLI were adequate, RMSEA was not within the recommended threshold. Still, one has to note that the confidence interval associated with RMSEA is within the range of acceptability and that this construct is measured with a relatively low number of items and a simply structure, which both contributes to the relatively low RMSEA.

Alpha for malleability
```{r}
variables %>%
  select(mall1:mall4) %>%
  alpha()
```

Omega total for malleability
```{r}
variables %>%
  select(mall1:mall4) %>%
  ci.reliability()
```

### Modern racism
```{r}
racism <- 'raci =~ rac1 + rac2 + rac3 + rac4 + rac5 + rac6 + rac7'
fit_racism <- cfa(racism, data = dehumi, estimator = 'MLR')
summary(fit_racism, fit.measures = TRUE, standardized = TRUE)
```
Once again, the unsatisfactory RMSEA can be attributed to the low number of items and the simple measurement structure.

Alpha for modern racism
```{r}
variables %>%
  select(rac1:rac7) %>%
  alpha()
```

Omega total for modern racism
```{r}
variables %>%
  select(rac1:rac7) %>%
  ci.reliability()
```

### Xenophobia
```{r}
xenoph <- 'xeno =~ xeno1 + xeno2 + xeno3 + xeno4 + xeno5'
fit_xenoph <- cfa(xenoph, data = dehumi, estimator = 'MLR')
summary(fit_xenoph, fit.measures = TRUE, standardized = TRUE)
```
Alpha for xenophobia
```{r}
variables %>%
  select(xeno1:xeno5) %>%
  alpha()
```
Omega total for xenophobia
```{r}
variables %>%
  select(xeno1:xeno5) %>%
  ci.reliability()
```
This scale is also satisfactory in terms of model fit.

Overall, the scales appear to have satisfactory factor structure as well as acceptable internal consistency and model-based composite reliabilities.

# Structural model
```{r}
final <- '
  tend =~ east + west + hung  
    east =~ syr + turk + lib
    west =~ ger + dane + usa
    hung =~ hun + hunw + hune 
  malleab =~ mall1 + mall2 + mall3 + mall4 
  racism =~ rac1 + rac2 + rac3 + rac4 + rac5 + rac6 + rac7
  xenoph =~ xeno1 + xeno2 + xeno3 + xeno4 + xeno5

  xenoph ~ malleab + tend
  racism ~ malleab + tend
  malleab ~ tend

  xenoph ~~ racism'

fit_model <- cfa(final, data = dehumi, estimator = 'MLR')
summary(fit_model, fit.measures = TRUE, standardized = TRUE)
```
General dehumanization tendency negatively predicted group malleability, modern racism, and xenophobia. Moreover, malleability positively predicted xenophobia and modern racism.
